---
title: "Lab Group 3: Project 1"
author: "Yinbo Zhao 520173668  Tierui Yang 520197398  Huiyu Ren 520192902 Tapaau Tualaulelei 520109243"
subtitle: "Project 1"
date: "University of Sydney | MATH1005 | Lab Class 22"
output:
  html_document:
    fig_caption: yes
    number_sections: yes
    self_contained: yes
    theme: flatly
    toc: true
    toc_depth: 3
    toc_float: true
    code_folding: hide
---

<br>

# Executive Summary
- The aim of this investigation is to determine the likeliness of developing diabetes mellitus for Pima Indian Mothers and non-Mothers of at least 21 years and older by observing some of its known causal factors such as Glucose and Insulin levels etc. 

- It has been reported that Pima Indians have the highest prevalence of non-insulin dependent diabetes mellitus world-wide which has reached 70% in the 55-64 year age-group (Charles et al., 1997). Though it is very common amongst many American Indian tribes, its genetic susceptibility within this population probably explains why it is an "*epidemic*" (Knowler et al., 1993). 

- For Pima Indian mothers, *serum insulin, body mass index, diabetes predigree function* and *age* are the major causal factors for developing diabetes as opposed to young Pima Indian females who are reportedly known to develop diabetes as an indication of the following causal factors: *age, serum insulin, body mass index, hereditary diabetes, diabetes pedigree, dietary habits etc* (Das, 2014). Both are seen to have hereditary implications of the disease and this has been a common determinant of Diabetes Mellitus for Pima Indians. However, a report by Knowler (1993) shows that though it is widely genetic within Pima Indians, the gradual increase of the incidence throughtout decades provides further evidence of environmental-genetic interaction meaning that some may be due to dietary habits. Hence why we are to observe the extent of these relationships between indicators of Diabetes (by medical standards) through the variables below. 

- The main discoveries are to: 
    - Identify the linkage between Glucose Plasma levels and Blood Pressure at different age ranges as well as the relationship between Glucose and Insulin. 
    - The likeliness of diabetes occurring hereditarily (*by observing the DiabetesPedigreeFunction*) as opposed to Class Outcome result or to observe whether there is a connectivity with respect to the data set.
    - Determine the linear regression line between these factors.
    
<br>

# Full Report

## Initial Data Analysis (IDA)

### Source

- The data was originally collected from the National Institute of Diabetes and Digestive and
Kidney Diseases survey on females of Pima Indian descent during the 1990's.

- The information was provided by Vincent Sigillito (*vg@alpcen.apl.jhu.edu*), Research center, RMI Leader, Applied Physics Lavoratory, The Johns Hopkins University, Johns Hopkins Road, Laurel, MD 20707 (301) 953-6231 

#### Description 

- The specified patients within this data frame are all females of *Pima Indian Heritage of at least 21 years of age*. The sample population, as depicted from the source, live near Phoenix, Arizona, USA and was received and finalized on the **9th of May, 1990**.

- The data frame has the illustrated  **768 rows** that indicates the total number of sample subjects, as well as **9 columns** which designates the 8 observed variables in relation to diabetes plus the binary-valued outcome variable.

- **Note:** The diagnostic binary-valued outcome variable as seen within the data set is to investigate whether the subjected patient does have symptoms of diabetes in accordance to the World Health Organisation criteria which states highlights *"if the 2 hour post-load plasma glucose was at least 200 mg/dl at any survey examination or if found during routine medical care.." *.(Smith et al., 1988) and by this attribute, we are given the 9th column where the outcome *1 means that person has diabetes* and *0 if they do not*.

#### Limitations and Reliability

- The main issue with the data set is that the **Class Outcome variable** does not specify whether the diabetic type is either Type 1 (*genetic disorder which appears at a very young age*) or Type 2 (*arises from dietary habits which develops overtime*) which can prove to be a posing confounding factor when drawing conclusions.

- However, the dataset is widely reliable for drawing statistical conclusions because this particular survey has been carried out every two years since 1953 (Knowler et al., 1993) as to observe any changes of this epidemic within the population of Pima Indians and has been used in many research.

#### Potential Stakeholders

- The possible **Stakeholders** for this research type are the Pima Indian population, institutional teachers, data architects and/or medical centers.

#### Format

The acquired data set has the following columns/variables:

<br>

  - **Pregnancies:** Number of times pregnant
  
  - **Glucose:** Plasma glucose concentration a 2 hours in an oral glucose tolerance test.
  
  - **BloodPressure:** Diastolic blood pressure (mm Hg)
  
  - **SkinThickness:** Triceps skin fold thickness (mm)
  
  - **Insulin:** 2-hour serum insulin (mu U/ml)
 
  - **BMI:** Body mass index (weight in kg/(height in m)^2)

  - **DiabetesPedigreeFunction:** Diabetes Pedigree Function
  
  - **Age:** Age in years
  
  - **Outcome:** Class variable (0 or 1, *where 1=positive*)
  
  <br>
  
However, the **key variables** that this report hopes to prioritize are: *Glucose, BloodPressure, Insulin and the Class Outcome* 

- The imported data frame shows:
```{r}
## read in data
f=read.csv("data/diabetes.csv", header= T)

## show classification of variables
dim(f)
str(f)
Glucose=f$Glucose
BloodPressure=f$BloodPressure

```

<br>

## Research Question 1: What is the relation between Glucose and BloodPressure

- One major causal factor of Diabetes by medical standards is *High Glucose levels* which then induces *High Blood Pressure* due to the rejection of insulin in their body.The Insulin hormone is used by the body to digest blood sugar and convert it into energy, but a person with Type 2 diabetes has this resistance to insulin which then causes the build up of sugar in the bloodstream (Canopy Health, 2018); Hence, the possible tendency of the patient to develop High blood pressure as well. 

- Stage 1 High Blood Pressure is when diastolic ranges from **80-89** whereas Stage 2 HBP has a range of **90 or higher**. However, for Glucose levels, it has been identified that those with the prospect of developing Diabetes show an Oral Glucose Tolerance Test (OGTT) of **140 to 199 mg/dL**  

*(descriptions for each analytic display have been provided below)* 
```{r}
## read data 
forigin = read.csv("data/diabetes.csv")
##remove value = 0 reasonably
f=data.frame(forigin)
f<-f[which(f$Glucose!=0),]
f<-f[which(f$BloodPressure!=0),]
f<-f[which(f$SkinThickness!=0),]
f<-f[which(f$Insulin!=0),]
f<-f[which(f$BMI!=0),]
f<-f[which(f$DiabetesPedigreeFunction!=0),]
f<-f[which(f$Age!=0),]
rownames(f)<-NULL
## show classification of variables
str(f)
Glucose = f$Glucose
BloodPressure = f$BloodPressure
```
<br>
**A) BloodPressure:**

- The Blood pressure of the subjects fall into the range between *60-80*, judging from the Graphical summaries. Further clarification can be observed from the Measures of Central Tendency and Measures of Variation below. 
```{r}
## write code here
par(mfrow = c(1,2))
hist(BloodPressure,xlab = "BloodPressure (mm Hg)")
boxplot(BloodPressure,horizontal = T,xlab = "BloodPressure (mm Hg)")
```

**1. The mean Value of BloodPressure**
```{r}
mean(BloodPressure)
```

**2. The standard deviation of BloodPressure**
```{r}
sd(BloodPressure)
```

**3. The median Value of BloodPressure**
```{r}
median(BloodPressure)
```

**4. The IQR Value of BloodPressure**
```{r}
IQR(BloodPressure)
```

**5. Normal Curve of BloodPressure:**
```{r}
curve(dnorm(x,round(mean(BloodPressure),0),round(sd(BloodPressure),0)),from=0,to=140,ylab="Density",main="Normal Curve of BloodPressure",xlab = "BloodPressure (mm Hg)")
```

<br>

Judging from the Curve above: 

- 68% people's Blood Pressure is in the range [58, 83]
```{r}
mean(BloodPressure)-sd(BloodPressure)
mean(BloodPressure)+sd(BloodPressure)
```

- 95% people's Blood Pressure is in the range [46, 96]
```{r}
mean(BloodPressure)-2*sd(BloodPressure)
mean(BloodPressure)+2*sd(BloodPressure)
```

- 99.7% people's Blood Pressure is in the range [33, 108], and  the rate of people with abnormal Blood Pressure (Hypertension- elevated Bloodpressure)shows:
```{r}
mean(BloodPressure)-3*sd(BloodPressure)
mean(BloodPressure)+3*sd(BloodPressure)
abbp=data.frame(BloodPressure)
abbp <- abbp[which(BloodPressure>=90),]
length(abbp)/length(BloodPressure)
```



<br>

***Analysis for BloodPressure:***

- On average, majority of the dataset has blood Pressure below 80 which is relatively low or normal with respect to the standard blood pressure of patients with diabetes (80 upwards) which can pose assumptions in this regard. Additionally, by observing the spread or variation of data, we can estimate that from the dataset, only about 0.3% deviate approximately 36 mm Hg from the mean of around 70 mm Hg.

- High blood pressure has always been one of the many indicators for diabetes- but more so in Type 2 as opposed to Type 1 patients, and to provide an idea of this number- its about 3/10 people with Type 1 diabetes and around 8/10 for those with Type 2 diabetes (Knott, 2020). This means that though someone may have diabetes, if they are Type 1 which is hereditary, then the likeliness of having High blood pressure is small. This can also be used to question why a large number of diabetic patients do not have High Blood Pressure or why one who does not have diabetes has high blood pressure (by observing the Class Outcome column) and this, in turn, can pose as a confounder in the observation of this variable. 

<br> 

To bring this finding into perspective, we observe the BloodPressure of patients with attribute 1 in Class Outcome (Diabetic patients) as opposed to patients without Diabetes ("0") as seen below: 

```{r}
Outcome = factor(f$Outcome, levels = c("0", "1"))
boxplot(BloodPressure ~ Outcome, horizontal=T, col=c("green", "red"), main = "Classification (Outcome attribute) by BloodPressure",xlab = "BloodPressure (mm Hg)")
```

- The above plot also indicates that though those who have Diabetes is higher than those who do not, the difference is very small and only a small variation exceeds 80 mm Hg. Thus why BloodPressure alone may or may not be a strong indicator of Diabetes Mettilus but this can vary depending on what type of diabetes each patient has since those with Type 2 have a greater tendency of High BloodPressure compared to those of Type 1 (Knott, 2020). 

<br>

**B) Glucose:**

```{r}
## write code here
par(mfrow = c(1,2))
hist(Glucose, xlab = "Glucose (mg/dL)")
boxplot(Glucose,horizontal = T, xlab = "Glucose (mg/dL)")
```

**1. The mean Value of Glucose**
```{r}
mean(Glucose)
```

**2. The standard deviation of Glucose**
```{r}
sd(Glucose)
```

**3. The median Value of Glucose**
```{r}
median(Glucose)
```

**4. The IQR Value of Glucose**
```{r}
IQR(Glucose)
```

**5. Normal Curve of Glucose:**
```{r}
curve(dnorm(x,round(mean(Glucose),0),round(sd(Glucose),0)),from=0,to=240,ylab="Density",main="Normal Curve of Glucose", xlab = "Glucose (mg/dL)")
```

<br>

From the curve above we can see that: 

- 68% people's Glucose is in the range [91, 153]
```{r}
mean(Glucose)-sd(Glucose)
mean(Glucose)+sd(Glucose)
```

- 95% people's Glucose is in the range [61, 184]
```{r}
mean(Glucose)-2*sd(Glucose)
mean(Glucose)+2*sd(Glucose)
```

- 99.7% people's Glucose is in the range [30, 215] and the rate of people with abnormal Glucose:
```{r}
mean(Glucose)-3*sd(Glucose)
mean(Glucose)+3*sd(Glucose)
abglucose=data.frame(Glucose)
abglucose <- abglucose[which(Glucose>=140),]
length(abglucose)/length(Glucose)
```



***Analysis for Glucose:***

- By calculating the mean of the Glucose variable, we are able to also depict that majority of the dataset subjects have Glucose levels that are below that of critical or glucose levels of those who have Diabetes in accordance to medical standards. To recall, diabetic patients have OGTT of 140-199 mg/dL whereas the data has a mean of about 122 mg/dL and a variation (sd) of about 31. On another note, by using the 68% 95% 99.7% rule, we can state that only about 0.3% of this dataset deviates approximately 31 mm/dL from the mean. 

- However, if we further investigate this attribute, we can observe the distribution of the patients with Diabetes and those without diabetes with respect to their spread of Glucose Plasma Levels. 

```{r}
Outcome = factor(f$Outcome, levels = c("0", "1"))
boxplot(Glucose ~ Outcome, horizontal=T, col=c("green", "red"), main = "Classification (Outcome attribute by Glucose ", xlab = "Glucose (mg/dL)")
```
<br>

- From the comparison shown above, we can say that the differences between the Glucose levels between Diabetic and Non-Diabetic patients are vast and though the mean value is relatively low, this can only be a reflection of the ***Simpson's Paradox*** which simply means that when grouped together, values of Non-Diabetics have a major impact on the aggregate of itself and those of Diabetic patients.

- The data above also proves that exceeding glucose levels is a major/main causal factor of Diabetes regardless of the Type.

<br>

### Glucose vs BloodPressure

**1. Scatter Plot:** 
```{r}
plot(Glucose, BloodPressure, xlab = "Glucose (mg/dL)",ylab = "BloodPressure (mm Hg)")
```

**Correlation coefficient**
```{r}
ybar=mean(BloodPressure)
y1=BloodPressure-mean(BloodPressure)
ySD=sd(BloodPressure)
ystd=y1/ySD
xbar=mean(Glucose)
x1=Glucose-mean(Glucose)
xSD=sd(Glucose)
xstd=x1/xSD
r=mean(ystd*xstd)
r
```

***Analysis for the scatter plot:***

Because the correlation coefficient is 0.2, this means that the correlation between glucose and blood pressure is not obvious, but we can still slightly observe from the figure that as glucose increases, more points are distributed in places where blood pressure is higher.

**2. HeatMap:**
```{r}
heatmap(table(BloodPressure, Glucose), Rowv = NA, Colv = NA, xlab = "Glucose (mg/dL)", ylab = "BloodPressure (mm Hg)")
```

<br>

***Analysis for the HeatMap:***

- After carefully observing the heat map, we found that the dark red color block appear higher as the value increases, which means that higher blood pressure will be accompanied by higher glucose.

**3. Linear model between glucose and BloodPressure:**

```{r}
lm(BloodPressure~Glucose)
```

Thus The equation of the regression line is

      BloodPressure = 60.23 + 0.085 * Glucose
      
      
```{r}
plot(Glucose, BloodPressure, xlab = "Glucose (mg/dL)",ylab = "BloodPressure (mm Hg)")
abline(lm(BloodPressure~Glucose))
```

***Analysis for the linear model:***

- It is not difficult to conclude from the figure that with the increase of glucose, blood pressure shows a slight upward trend.

**4. Diagram of the Residual**
```{r}
l = lm( BloodPressure ~ Glucose)
plot(Glucose, l$residuals,xlab = "Glucose (mg/dL)")
abline(h=0, col = "red")
```

***Analysis for the residual diagram:***

- From the residual diagram, we can see that the points we plot around the line with residual equal to 0 are randomly scattered up and down, indicating that the regression line fits the original observations well.

### Analysis for this relationship between Glucose and BloodPressure

- The correlation between these two variables as calculated from the Scatterplot and linear model tells us that there is indeed a positive correlation between the two-  though small;This simply indicates that when the BloodPressure increases along the axis, the Glucose level also increases.   

- The same can be seen from the HeatMap data representation where the higher blood pressure will be accompanied by higher glucose.

- Therefore, we can infer that with the increase of glucose content in the human body, there will be a slight upward trend in human blood pressure- regardless of the diabetic type. This may be because when the blood glucose in the human body increases, the blood will become more viscous and the blood flow with more resistance, resulting in increased blood pressure. Meanwhile, the residual diagram also confirms that our model are reasonable.


<br>

## Research Question 2: What is the relation between Insulin and Glucose

- As explained earlier, Insulin is a hormone that helps our body digest the blood sugar in order to produce energy, However, for a diabetic patient, lack of insulin or lack of insulin usage is its main causal factor for sugar accumulation in the bloodstream, thus- Diabetes. Insulin shortage can be dangerous for both Types of Diabetes and by this information alone, we can hypothesize the relationship between Insulin and Glucose.  

- In accordance to medical observations, those with insulin levels higher than **276 mu U/ml** after fasting are the most susceptible to developing Diabetes (Lillioja et al., 1993).  

```{r}
## read data 
forigin = read.csv("data/diabetes.csv")
##remove value = 0 reasonably
f=data.frame(forigin)
f<-f[which(forigin$Glucose!=0),]
f<-f[which(f$BloodPressure!=0),]
f<-f[which(f$SkinThickness!=0),]
f<-f[which(f$Insulin!=0),]
f<-f[which(f$BMI!=0),]
f<-f[which(f$DiabetesPedigreeFunction!=0),]
f<-f[which(f$Age!=0),]
rownames(f)<-NULL
## show classification of variables
str(f)
Insulin = f$Insulin
Glucose = f$Glucose
Outcome = factor(f$Outcome, levels = c("0", "1"))
```

***A) Insulin:***
<br>

From the graphical summaries below, we can see that the Insulin range is from 0-200 (Histogram) but further observations of the Measures of Spread and Central Tendency shows the following:
```{r}
## write code here
par(mfrow = c(1,2))
hist(Insulin, xlab = "Insulin(mu U/ml)")
boxplot(Insulin,horizontal = T, xlab = "Insulin(mu U/ml)")
```

**1. The mean Value of Insulin**
```{r}
mean(Insulin)
```

**2. The standard deviation of Insulin**
```{r}
sd(Insulin)
```

**3. The median Value of Insulin**
```{r}
median(Insulin)
```

**4. The IQR Value of Insulin** 
```{r}
IQR(Insulin)
```

**5. the rate of people with abnormal Insulin**
```{r}
abins=data.frame(Insulin)
abins <- abins[which(Insulin>=150),]
length(abins)/length(Insulin)
```
***Analysis for Insulin:*** 

- The mean obtained from the dataset tells us that majority of the samples have insulin levels around **156 mu u/ml**, which by standard is lower than the number for those who have diabetes. This then proves statement by *M. Regina Castro (2020)* that High Insulin tests alone cannot mean Diabetes. Again, the spread shows us that from the mean, only about 0.3% deviates about **119 mu U/ml** from the mean of about **156 mu U/ml**. 

- However, further observation into this attribute between Diabetic and Non-Diabetic shows that:

```{r}

boxplot(Insulin ~ Outcome, horizontal = T, col= c("green", "red"), main = "Classification (Outcome attribute) by Insulin", xlab = "Insulin(mu U/ml)")
```

<br>

- There is a variation between the two classifications when it comes to Insulin levels, thus proving that exceeding levels of Insulin also induces Glucose accumulation in the bloodstream which in turn contributes to Diabetes development.

- However, it is important to also note that for both grouped data, their maximum values were still relatively smaller or lower than the standard degree of diabetic implication through the insulin serum test which can mean one of two things: either majority of the diabetic patients had fasted for more than 2 hours (before testing) which led to significant lowering of insulin levels throughout the Insulin serum test or either majority are Type 1 diabetes which has a lower likeliness of high Insulin levels (Watson, 2021).  


***B) Glucose:***

```{r}
## write code here
par(mfrow = c(1,2))
hist(Glucose, xlab = "Glucose (mg/dL)")
boxplot(Glucose,horizontal = T, xlab = "Glucose (mg/dL)")
```

**1. The mean Value of Glucose**
```{r}
mean(Glucose)
```

**2. The standard deviation of Glucose**
```{r}
sd(Glucose)
```

**3. The median Value of Glucose**
```{r}
median(Glucose)
```

**4. The IQR Value of Glucose**
```{r}
IQR(Glucose)
```

**5. Normal Curve of Glucose:** 
```{r}
curve(dnorm(x,round(mean(Glucose),0),round(sd(Glucose),0)),from=0,to=240,ylab="Density",main="Normal Curve of Glucose", xlab = "Glucose (mg/dL)")
```

68% people's Glucose is in the range [91, 153]
```{r}
mean(Glucose)-sd(Glucose)
mean(Glucose)+sd(Glucose)
```

95% people's Glucose is in the range [61, 184]
```{r}
mean(Glucose)-2*sd(Glucose)
mean(Glucose)+2*sd(Glucose)
```

99.7% people's Glucose is in the range [30, 215]
```{r}
mean(Glucose)-3*sd(Glucose)
mean(Glucose)+3*sd(Glucose)
```

the rate of people with abnormal Glucose
```{r}
abglucose=data.frame(Glucose)
abglucose <- abglucose[which(Glucose>=140),]
length(abglucose)/length(Glucose)
```


### Insulin vs Glucose

**1. Scatter Plot:**
```{r}
plot(Glucose,Insulin, xlab = "Glucose (mg/dL)", ylab = "Insulin(mu U/ml)")
```

**Correlation coefficient**
```{r}
xbar=mean(Glucose)
x1=Glucose-mean(Glucose)
xSD=sd(Glucose)
xstd=x1/xSD
ybar=mean(Insulin)
y1=Insulin-mean(Insulin)
ySD=sd(Insulin)
ystd=y1/ySD
r=mean(xstd*ystd)
r
```

***Analysis for the scatter plot:***

- From the scatter plot, we can clearly find that as the glucose rises, more points are distributed in the places where the insulin of the human body is higher. This proves the individual observations of the variable as well as medical understanding that Glucose Plasma Levels are directly proportional to Insulin levels. 

**2. HeatMap:**
```{r}
heatmap(table(Insulin,Glucose), Rowv = NA, Colv = NA, xlab = " Glucose (mg/dL)" , ylab = "Insulin(mu U/ml)")
```

***Analysis for the HeatMap:***

- From the heatmap we can observe that as the glucose value increases, more color blocks are distributed where the insulin value is also higher.

**3. Linear model between ins and glucose:**

```{r}
lm(Insulin~Glucose)
```

Thus The equation of the regression line is

      Glucose = -118.41 + 2.24 * BloodPressure
      
      
```{r}
plot(Glucose,Insulin, xlab = "Glucose (mg/dL)", ylab = "Insulin(mu U/ml)")
abline(lm(Insulin~Glucose))
```

***Analysis for the linear model:***

- From the regression line we drew, the insulin content in the human body will increase with the increase of glucose. Therefore they are directly proportional and have a positive correlation.

**4. Diagram of the Residual**
```{r}
l = lm( Insulin ~ Glucose)
plot(Glucose, l$residuals, xlab = "Glucose (mg/dL)")
abline(h=0, col = "red")
```

***Analysis for the residual diagram:***

- From the residual diagram, we can see that the points we plot around the line with residual equal to 0 are randomly scattered up and down, indicating that the regression line fits the original observations well.

<br>

### Analysis for this relationship between Insulin and Glucose

- From the above figures, we can easily find that when the glucose level in the human body increases, the insulin level will also increase. This may be because the main function of insulin is to control the level of blood in the blood after eating. It could prevents glucose from staying in the bloodstream and causes blood glucose to rise by telling cells to absorb glucose as it comes in. 

- Again, it shows connectivity between genetic susceptibility of Diabetes to appear in Pima Indians, even when compared to other races such as Causassions of similar degrees of glucose tolerance as observed in a data analysis written by Charles et al.,(1997) who concluded that Pima Indians are highly insulin-resistant thus induces high glucose accumulation in their bloodstream.  

## Conslusion

In culmination, we can state that by observing the following variables: *Glucose, BloodPressure, Insulin* and the *Outcome* Attribute of Pima Indian females of above 21 years of age, we can also infer that: 

- ***BloodPressure*** in this regard is a more unlikely indicator of a person having Diabetes- especially with this dataset. However, because we cannot identify whether a person has Type 1 or Type 2 diabetes mellitus, we can assume that this conclusion means that majority who were diagnoses were of Type 1 diabetes. 

- ***Glucose*** is the main or a major attribute to Diabetes regardless of the Type classification. It was through this singular observation that we were able to see the clear variation between females with Diabetes and those who do not have Diabetes and when plotted against variables- Insulin and BloodPressure, there was a clear positive correlation though it was small for BloodPressure. 

- ***Insulin*** has been defined to go hand in hand with Glucose levels since through medical observations, its the tendency of a person's system to resist the usage of Insulin for sugar breakdown is the what causes Diabetes for both Type 1 and 2. 

Through this investigation, we were also able to:

- Determine the line of regression as well as the correlation coefficient between the observed attributes where both indications showed positive correlations and had proved that each variable is directly proportional to the other. 

- Though Pima Indians are known to have a higher prevalence of Diabetes throughout generations- widely hereditary, we were able to see an indication of that throughout the dataset such as BloodPressure levels being relative lower in the Diabetic group- hence proving that those with Type 1 diabetes (genetics) have a lower prospect of having BloodPressure (Knott, 2020) as opposed to those with Type 2 Diabetes. Additionally, insulin levels of Type 1 diabetics has to do more with the less production of insulin (lower insulin levels) rather than high insulin levels as seen in Type 2 patients (Watson, 2021) as depicted from the data above. 

# List of References:

- Canopy Health. (2018). Understanding the Relationship between High bloodPressure and Type 2 Diabetes. *Health and Wellness*, https://www.canopyhealth.com/understanding-the-relationship-between-high-blood-pressure-and-type-2-diabetes.  

- Charles, M. A., Eshwege, E., Bennett, P. H. (1997). Non-insulin-dependent diabetes in populations at risk: the Pima Indians. *Diabetes & Metabolism, 23Suppl 4*, 6-9. 

- Das, R., N. (2014). Determinants of Diabetes Mellitus in the Pima Indian Mothers and Indian Medical Students. *The Open Diabetes Journal*, *7*(1), 5-13. https://doi.org/10.2174/187652460140701005 

- Knowler, W. C., Nelson, R. G., Saad, M. F., Bennt, P. H., Pettitt, D. J.(1993). Determinants of Diabetes Mellitus in the Pima Indians. *Diabetes Care*, *16*(1), 216-227. https://doi.org/10.2337/diacare.16.1.216 

- Lillioja, S., Mott, D. M., Spraul, M., Ferraro, R., Foley, J. E., Knowler, W. C. ...Bogardus, C. (1993). Insulin Resistance and Insulin Secretory Dysfunction as Precursors of Non-Insulin-Dependent Diabetes Mellitus:           Prospect Studies of Pima Indian. *The New England Journal of Medicine*, *329*(27), 1988-1992. https://doi.org/10.1056/NEJM199312303292703 

- Mayo Clinic. (Dec 2020). Hyperinsulinemia: Is it diabetes?. https://protect-au.mimecast.com/s/vJdLCmO5glu5MBN5mSGd_EW?domain=mayoclinic.org 

- Smith, J. W., Evenhart, J. E., Dickson, W. C., Knowler, W. C., Johannes, R. S., (1998). Using the ADAP Learning Algorithm tp Forecast the Onset of Diabetes Mellitus.*Proceedings of the Annual Symposium on Computer              Application in Medical Care*, 261-265.

- Watson, S. (March 2021). How Insulin Treatment Lowers your blood sugar. *Diabetes Treatment Home*, https://protect-au.mimecast.com/s/TNZvCP7LAXf468RQYhzSfTq?domain=webmd.com  




