---
title: "Lab Group 3: Project 1"
author: "Yinbo Zhao 520173668  Tierui Yang 520197398  Huiyu Ren 520192902 Tapaau Tualaulelei 520109243"
subtitle: "Project 1"
date: "University of Sydney | MATH1005 | Lab Class 22"
output:
  html_document:
    fig_caption: yes
    number_sections: yes
    self_contained: yes
    theme: flatly
    toc: true
    toc_depth: 3
    toc_float: true
    code_folding: hide
---

<br>

# Executive Summary
- The aim of this investigation is to determine the likeliness of developing diabetes mellitus for Pima Indian Mothers and non-Mothers of at least 21 years and older by observing some of its known causal factors such as Glucose and Insulin levels etc. 

- It has been reported that Pima Indians have the highest prevalence of non-insulin dependent diabetes mellitus world-wide which has reached 70% in the 55-64 year age-group (Charles et al., 1997). Though it is very common amongst many American Indian tribes, its genetic susceptibility within this population probably explains why it is an "*epidemic*" (Knowler et al., 1993). 

- For Pima Indian mothers, *serum insulin, body mass index, diabetes predigree function* and *age* are the major causal factors for developing diabetes as opposed to young Pima Indian females who are reportedly known to develop diabetes as an indication of the following causal factors: *age, serum insulin, body mass index, hereditary diabetes, diabetes pedigree, dietary habits etc* (Das, 2014). Both are seen to have hereditary implications of the disease and this has been a common determinant of Diabetes Mellitus for Pima Indians. However, a report by Knowler (1993) shows that though it is widely genetic within Pima Indians, the gradual increase of the incidence throughtout decades provides further evidence of environmental-genetic interaction meaning that some may be due to dietary habits. Hence why we are to observe the extent of these relationships between indicators of Diabetes (by medical standards) through the variables below. 

- The main discoveries are to: 
    - Identify the linkage between Glucose Plasma levels and Blood Pressure at different age ranges as well as the relationship between Glucose and Insulin. 
    - The likeliness of diabetes occurring hereditarily (*by observing the DiabetesPedigreeFunction*) as opposed to Class Outcome result or to observe whether there is a connectivity with respect to the data set.
    - Determine the linear regression line between these factors.
    
<br>

# Full Report

## Initial Data Analysis (IDA)

### Source

- The data was originally collected from the National Institute of Diabetes and Digestive and
Kidney Diseases survey on females of Pima Indian descent during the 1990's.

- The information was provided by Vincent Sigillito (*vg@alpcen.apl.jhu.edu*), Research center, RMI Leader, Applied Physics Lavoratory, The Johns Hopkins University, Johns Hopkins Road, Laurel, MD 20707 (301) 953-6231 

#### Description 

- The specified patients within this data frame are all females of *Pima Indian Heritage of at least 21 years of age*. The sample population, as depicted from the source, live near Phoenix, Arizona, USA and was received and finalized on the **9th of May, 1990**.

- The data frame has the illustrated  **768 rows** that indicates the total number of sample subjects, as well as **9 columns** which designates the 8 observed variables in relation to diabetes plus the binary-valued outcome variable.

- **Note:** The diagnostic binary-valued outcome variable as seen within the data set is to investigate whether the subjected patient does have symptoms of diabetes in accordance to the World Health Organisation criteria which states highlights *"if the 2 hour post-load plasma glucose was at least 200 mg/dl at any survey examination or if found during routine medical care.." *.(Smith et al., 1988) and by this attribute, we are given the 9th column where the outcome *1 means that person has diabetes* and *0 if they do not*.

#### Limitations and Reliability

- The main issue with the data set is that the **Class Outcome variable** does not specify whether the diabetic type is either Type 1 (*genetic disorder which appears at a very young age*) or Type 2 (*arises from dietary habits which develops overtime*) which can prove to be a posing confounding factor when drawing conclusions.

- However, the dataset is widely reliable for drawing statistical conclusions because this particular survey has been carried out every two years since 1953 (Knowler et al., 1993) as to observe any changes of this epidemic within the population of Pima Indians and has been used in many research.

#### Potential Stakeholders

- The possible **Stakeholders** for this research type are the Pima Indian population, institutional teachers, data architects and/or medical centers.

#### Format

The acquired data set has the following columns/variables:

<br>

  - **Pregnancies:** Number of times pregnant
  
  - **Glucose:** Plasma glucose concentration a 2 hours in an oral glucose tolerance test.
  
  - **BloodPressure:** Diastolic blood pressure (mm Hg)
  
  - **SkinThickness:** Triceps skin fold thickness (mm)
  
  - **Insulin:** 2-hour serum insulin (mu U/ml)
 
  - **BMI:** Body mass index (weight in kg/(height in m)^2)

  - **DiabetesPedigreeFunction:** Diabetes Pedigree Function
  
  - **Age:** Age in years
  
  - **Outcome:** Class variable (0 or 1, *where 1=positive*)
  
  <br>
  
However, the **key variables** that this report hopes to prioritize are: *Glucose, BloodPressure, Insulin and the Class Outcome* 

- The imported data frame shows:
```{r}
## read in data
f=read.csv("data/diabetes.csv", header= T)

## show classification of variables
dim(f)
str(f)
Glucose=f$Glucose
BloodPressure=f$BloodPressure

```

<br>

## Research Question 1: What is the relation between Glucose and BloodPressure

- One major causal factor of Diabetes by medical standards is *High Glucose levels* which then induces *High Blood Pressure* due to the rejection of insulin in their body.The Insulin hormone is used by the body to digest blood sugar and convert it into energy, but a person with Type 2 diabetes has this resistance to insulin which then causes the build up of sugar in the bloodstream (Canopy Health, 2018); Hence, the possible tendency of the patient to develop High blood pressure as well. 

- Stage 1 High Blood Pressure is when diastolic ranges from **80-89** whereas Stage 2 HBP has a range of **90 or higher**. However, for Glucose levels, it has been identified that those with the prospect of developing Diabetes show an Oral Glucose Tolerance Test (OGTT) of **140 to 199 mg/dL**  

*(descriptions for each analytic display have been provided below)* 
```{r}
## read data 
forigin = read.csv("data/diabetes.csv")
##remove value = 0 reasonably
f=data.frame(forigin)
f<-f[which(f$Glucose!=0),]
f<-f[which(f$BloodPressure!=0),]
f<-f[which(f$SkinThickness!=0),]
f<-f[which(f$Insulin!=0),]
f<-f[which(f$BMI!=0),]
f<-f[which(f$DiabetesPedigreeFunction!=0),]
f<-f[which(f$Age!=0),]
rownames(f)<-NULL
## show classification of variables
str(f)
Glucose = f$Glucose
BloodPressure = f$BloodPressure
```
<br>
**A) BloodPressure:**

- The Blood pressure of the subjects fall into the range between *60-80*, judging from the Graphical summaries. Further clarification can be observed from the Measures of Central Tendency and Measures of Variation below. 
```{r}
## write code here
par(mfrow = c(1,2))
hist(BloodPressure,xlab = "BloodPressure (mm Hg)")
boxplot(BloodPressure,horizontal = T,xlab = "BloodPressure (mm Hg)")
```
