---
title: "Relation among Glucose , Insulin and BloodPressure"
author: "Yinbo Zhao 520173668"
subtitle: "Project 1"
date: "University of Sydney | MATH1005 | Lab Class 22"
output:
  html_document:
    fig_caption: yes
    number_sections: yes
    self_contained: yes
    theme: flatly
    toc: true
    toc_depth: 3
    toc_float: true
    code_folding: hide
---

<br>

# Executive Summary
- The aim of this report is to ...
- The main discoveries are ...

<br>

# Full Report

## Initial Data Analysis (IDA)

- The data came from ...
- The data is/is not valid because ...
- Possible issues include ...
- Potential stakeholders include...
- Each row represents ...
- Each column represents ...
- The key variables are...

```{r}
## read in data
f = read.csv("data/diabetes.csv")

## show classification of variables
str(f)
glucose = f$Glucose
bp = f$BloodPressure

```


<br>

## Research Question 1: What is the relation between Glucose and BloodPressure

Insert text and analysis. 


BloodPressure:
```{r}
## write code here



par(mfrow = c(1,3))
hist(bp)
boxplot(bp)

```

The mean Value of BloodPressure
```{r}
mean(bp,na.rm = T)
```

The standard deviation of BloodPressure
```{r}
sd(bp,na.rm = T)
```

The median Value of BloodPressure
```{r}
median(bp,na.rm = T)
```

The IQR Value of BloodPressure
```{r}
IQR(bp,na.rm = T)
```

Normal Curve of BloodPressure:
```{r}
curve(dnorm(x,69,19),from=0,to=138,ylab="Density",main="N(69,361)")

```
68% people's BloodPressure is in the range [50, 88]
```{r}
qnorm(0.16,69,19)
qnorm(0.84,69,19)
```
95% people's BloodPressure is in the range [31, 107]
```{r}
qnorm(0.025,69,19)
qnorm(0.975,69,19)
```
99.7% people's BloodPressure is in the range [12, 126]
```{r}
qnorm(0.0015,69,19)
qnorm(0.9985,69,19)
```
Glucose:

```{r}
## write code here
par(mfrow = c(1,2))
hist(glucose)
boxplot(glucose)
```

The mean Value of Glucose
```{r}
mean(glucose,na.rm = T)
```

The standard deviation of Glucose
```{r}
sd(glucose,na.rm = T)
```

The median Value of Glucose
```{r}
median(glucose,na.rm = T)
```

The IQR Value of Glucose
```{r}
IQR(glucose,na.rm = T)
```
Normal Curve of Glucose:
```{r}
curve(dnorm(x,121,32),from=40,to=200,ylab="Density",main="N(121,1024)")

```
68% people's Glucose is in the range [89, 153]
```{r}
qnorm(0.16,121,32)
qnorm(0.84,121,32)
```
95% people's Glucose is in the range [57, 185]
```{r}
qnorm(0.025,121,32)
qnorm(0.975,121,32)
```
99.7% people's Glucose is in the range [25, 217]
```{r}
qnorm(0.0015,121,32)
qnorm(0.9985,121,32)
```

Glucose and BloodPressure

Scatter Plot:
```{r}
plot(bp, glucose)
```
HeatMap:
```{r}
heatmap(table(bp, glucose), Rowv = NA, Colv = NA)
```
Linear model between glucose and BloodPressure:

```{r}
lm( bp ~ glucose)
```
Thus The equation of the regression line is

      BloodPressure = 57.94 + 0.09 * Glucose
      
      
```{r}
plot(bp, glucose)
abline(lm(bp~glucose))

```

Diagram of the Residual
```{r}

l = lm( bp ~ glucose)
plot(bp, l$residuals)
abline(h=0, col = "red")

```




<br>

## Research Question 2: write research question here

Insert text and analysis. 

```{r}
## write code here

```


<br>

# References

Put references here in APA format.
